---
title: "USMI BR55"
output: html_notebook
---


```{r}
# Load main functions...
source("main.R", echo = F)
```

```{r}
# Source files location
source_dir <- "../datas/COHORTE BR55 PRINCIPALE"
# List recursively all files in sub filders, containing Graph.txt in their name
files <- list.files(source_dir, recursive = T, full.names = T, pattern = "Graph.txt")
```



### Data processing
```{r}
# Create a list of dataframes
dataframes_list <- lapply(files, parse_echo_graphs, test_cohorte = T)

# Concatenate the list of frames into one big frame
df <- data.frame()
for(i in 1:length(dataframes_list)){
  df <- dplyr::bind_rows(df,dataframes_list[[i]] )
}
# Remove the useless var
rm("dataframes_list")

#Convert the data.frame object to data.table object
setDT(df)

# Convert as numeric de mouse id value
df[, mouse_id := as.numeric(mouse_id)]

# Correct values for mouse 536 J7, measurements where performed 1 minute earlier
# Based on observed values, natural decreasing of the signal shouldn't make a huge difference
df[mouse_id == 536 & day == "J7" & msec > 400000, msec := msec + 60000]
   


# Check if no NAs introduced by numerical casting
stopifnot(sum(is.na(df[, mouse_id])) == 0)

# Open anapath data
anapath_data <- data.table::fread("../datas/COHORTE PRINCIPALE ANAPATH.csv")
   
# Join the two dataset by mouse id key   
joinned_datas <- df[anapath_data, on=c(mouse_id = "souris")]


# Load groups informations
Groups <- data.table::fread("mice_groups.csv")

# Set keys for automatic joint
setkey(Groups, Name)
setkey(joinned_datas, mouse_id)
joinned_final <- joinned_datas[Groups,]


# Open and read data for poid, volume and surfaces
datas_pvs <- data.table::fread("../datas/data_souris_poids_volume_surfaces.csv", dec = ",")

table_len <- nrow(datas_pvs)

# Convert wide format for poids and volume to wide
datas_pvs_long <- rbind(datas_pvs, datas_pvs, datas_pvs, datas_pvs)
datas_pvs_long$day <- c(rep("J0", table_len), rep("J3", table_len), rep("J7", table_len), rep("J11", table_len))
datas_pvs_long$poids <- c(datas_pvs$poids_J0, datas_pvs$poids_J3, datas_pvs$poids_J7, datas_pvs$poids_J11)
datas_pvs_long$volume <- c(datas_pvs$volume_J0, datas_pvs$volume_J3, datas_pvs$volume_J7, datas_pvs$volume_J11)

# Remove previous unused columns
datas_pvs_long <- datas_pvs_long[, -(3:10)]


# Merge thoses datas with joinned final
joinned_final_t <- merge(joinned_final, datas_pvs_long, by.x = c("mouse_id", "day"), by.y = c("souris", "day"))


# Compute late signals intensity
avg_signals <- c("mean_corrected", "mean_init", "mean_base")
signals_intensity <- df[msec > 600000, .(mean_corrected = adjusted_mean(.SD)[1], mean_init = adjusted_mean(.SD)[2], mean_base = adjusted_mean(.SD)[3]), by = .(mouse_id, day), .SDcols = c("T", "msec")]
# write.csv(signals_intensity, "../output_data/bound_signals_test_12082021.csv")


signals_intensity <- merge(signals_intensity, Groups, by.x = "mouse_id", by.y = "Name")


rm("anapath_data")
rm("datas_pvs")
rm("df")
rm("Groups")
saveRDS(df, "./processed_datas/df.rds")
saveRDS(signals_intensity, "./processed_datas/signals_intensity.rds")
saveRDS(datas_pvs_long, "./processed_datas/datas_pvs_long.rds")
saveRDS(joinned_final, "./processed_datas/joinned_final.rds")

```
### Read the datas
The rest of the script can be launched from this cell if full raw data are not available
Ask the data to the author in order to run the script properly

```{r}
source("main.R", echo = F)
df <- readRDS("./processed_datas/df.rds")
signals_intensity <- readRDS("./processed_datas/signals_intensity.rds")
datas_pvs_long <- readRDS("./processed_datas/datas_pvs_long.rds")
joinned_final <- readRDS("./processed_datas/joinned_final.rds")

```

```{r}
#Phase tardive
tardive_joinned <- joinned_final[msec > 600000, ]
tardive_joinned <- tardive_joinned[, Time := msec / 1000]

tardive_joinned <- tardive_joinned[, Groupe := ifelse(Group == "placebo", "Placebo", ifelse(Group == "demi", "axi 7.5", "axi 15"))]


tardive_joinned <- tardive_joinned[, enc_groups := ifelse(Group == "placebo", 1, ifelse(Group == "demi", 2, 3))] 


# Create output folder if necessairy
if(!dir.exists("../output_plots/")) dir.create("../output_plots/")
```
### Signals graphs for J11
```{r}
# For J11
g <- plot_graph_signals(tardive = tardive_joinned,
                        sel_day = "J11",
                        saving_location = "../output_plots/J11_late_signals.svg")


g
```
### Signals graphs for J0
```{r}
# For J0
g <- plot_graph_signals(tardive = tardive_joinned,
                        sel_day = "J0",
                        saving_location = "../output_plots/J0_late_signals.svg")


g
```

### Signals graphs for J3
```{r}
# For J3
g <- plot_graph_signals(tardive = tardive_joinned,
                        sel_day = "J3",
                        saving_location = "../output_plots/J3_late_signals.svg")


g
```


### Signals graphs for J7
```{r}
# For J7
g <- plot_graph_signals(tardive = tardive_joinned,
                        sel_day = "J7",
                        saving_location = "../output_plots/J7_late_signals.svg")


g
```


### Compute the Signal intensity for the late phase
```{r}
g <- ggplot(signals_intensity, aes(x = mean_base, y = mean_init))+
  geom_point(aes(col = Group))+
  theme_minimal() +
  facet_wrap(day ~ .)
g

```


### Correlation plots
```{r}
precoce <- joinned_final[msec < 500000, .(peak_intensity = unique(peak_intensity),area = unique(area),Group = unique(Group), time_to_peak = unique(time_to_peak), area_wash_in = unique(area_wash_in), area_wash_out = unique(area_wash_out), mean_transit_time = unique(mean_transit_time)), by = .(mouse_id, day)]


precoce <- merge(precoce,signals_intensity, by.x = c("mouse_id", "day"), by.y = c("mouse_id", "day"))

precoce$day <- factor(precoce$day, levels=c("J0", "J3", "J7", "J11"), labels=c("J0", "J3", "J7", "J11"))


precoce <- precoce[, Groupe := ifelse(Group.x == "placebo", "Placebo", ifelse(Group.x == "demi", "axi 7.5", "axi 15"))]

precoce$Groupe <- factor(precoce$Groupe, levels=c("Placebo", "axi 7.5", "axi 15"), labels=c("Placebo", "axi 7.5", "axi 15"))

require(corrplot)

M <- cor(precoce[, .(peak_intensity, area, time_to_peak, area_wash_in, area_wash_out, mean_corrected, mean_base, mean_init)])
corrplot(M, method="number")

```

# Peak intensity
```{r}
# g <- plot_variable_per_day_and_groups(precoce,
#                                       "peak_intensity",
#                                       "Peak intensity (a.u)",
#                                       "../output_plots/peak_intensity.svg")
# g

g <- variables_comparison_plots(precoce, "peak_intensity", "Peak intensity (a.u)")
g
ggsave(file="../output_plots/peak_intensity.svg", plot=g, width=10, height=8)
```

# Time to peak
```{r}
# g <- plot_variable_per_day_and_groups(precoce,
#                                       "time_to_peak",
#                                       "Time to peak (s)",
#                                       "../output_plots/time_to_peak.svg")
# g

g <- variables_comparison_plots(precoce, "time_to_peak", "Time to peak (s)")
g
ggsave(file="../output_plots/time_to_peak.svg", plot=g, width=10, height=8)
```

# Mean transit time
```{r}
# g <- plot_variable_per_day_and_groups(precoce,
#                                       "mean_transit_time",
#                                       "Mean transit time (s)",
#                                       "../output_plots/mean_transit_time.svg")
# g

g <- variables_comparison_plots(precoce, "mean_transit_time", "Mean transition time (s)")
g
ggsave(file="../output_plots/mean_transit_time.svg", plot=g, width=10, height=8)
```


# Area
```{r}
# g <- plot_variable_per_day_and_groups(precoce,
#                                       "area",
#                                       "Area (a.u)",
#                                       "../output_plots/area.svg")
# g

g <- variables_comparison_plots(precoce, "area", "Area under the curve (a.u)")
g
ggsave(file="../output_plots/area_under_curve.svg", plot=g, width=10, height=8)
```

# Area wash in 
```{r}
# g <- plot_variable_per_day_and_groups(precoce,
#                                       "area_wash_in",
#                                       "Area wash in (a.u*s)",
#                                       "../output_plots/area_wash_in.svg")
# g


g <- variables_comparison_plots(precoce, "area_wash_in", "Area wash in (a.u*s)")
g
ggsave(file="../output_plots/area_wash_in.svg", plot=g, width=10, height=8)

```

# Area wash out
```{r}
# g <- plot_variable_per_day_and_groups(precoce,
#                                       "area_wash_out",
#                                       "Area wash out (a.u*s)",
#                                       "../output_plots/area_wash_out.svg")
# g

g <- variables_comparison_plots(precoce, "area_wash_out", "Area wash out (a.u*s)")
g
ggsave(file="../area_wash_out.svg", plot=g, width=10, height=8)
```

# Mean 600s signal
```{r}
# g <- plot_variable_per_day_and_groups(precoce,
#                                       "mean_corrected",
#                                       "Mean 600s signal (a.u)",
#                                       "../output_plots/mean_600_signal.svg")
# g
g <- variables_comparison_plots(precoce, "mean_corrected", "MBs bound signal (a.u)")
g

ggsave(file="../output_plots/MBs_bound_signal.svg", plot=g, width=10, height=8)
```



```{r}
# Perform statistical tests... And sum them up in a dataframe
boxplots_tests <- function(selected_variable, df){
   groups_names <- c("Placebo", "axi 7.5", "axi 15")
   days <- c("J0", "J3", "J7", "J11")
  final_df <- data.table()
  for(aday in days){
    # Subset per day
    subseting <- df[day == aday,]
    
    # Perform shapiro test on each group per day
    rez_shapiro <- sapply(groups_names, FUN = function(x){
      shapiro.test(subseting[Groupe == x, ][[ selected_variable]])$p.value
      #ggqqplot(subseting[Groupe == x,][[selected_variable]])
    })
    
    print(paste0(selected_variable, " ~ Groupe"))
    form <- as.formula(paste0(selected_variable, " ~ Groupe"))

    rez_bartlett <- bartlett.test(form, data = subseting)$p.value
    
    # Anova
    
    anovas <- aov(form, data = subseting)
      
    # Kruskal
    kruskal <- kruskal.test(form, data = subseting)
    
    shapiro_boolean <- rez_shapiro <= 0.05
    bartlett_boolean <- rez_bartlett <= 0.05
    
    # Student Placebo to axi 7.5
    student_1 <- t.test(subseting[Groupe == "Placebo",][[ selected_variable]], subseting[Groupe == "axi 7.5",][[ selected_variable]])$p.value
   
    # Student Placebo to axi 15
    student_2 <- t.test(subseting[Groupe == "Placebo",][[ selected_variable]], subseting[Groupe == "axi 15",][[ selected_variable]])$p.value
    
    # Student axi 7.5 to axi 15
    student_3 <- t.test(subseting[Groupe == "axi 7.5",][[ selected_variable]], subseting[Groupe == "axi 15",][[ selected_variable]])$p.value
    
    
    tmp <- data.table(
              var = selected_variable,
              day = aday,
              shapiro_placebo = rez_shapiro[1],
              shapiro_axi_7 = rez_shapiro[2],
              shapiro_axi_15 = rez_shapiro[3],
              bartlett = rez_bartlett,
              kruskal_pval = kruskal$p.value,
              anova_Pr = ifelse((sum(shapiro_boolean) == length(unique(subseting$Groupe))) & bartlett_boolean, summary(anovas)[[1]]$`Pr(>F)`[[1]], NA),
              student_placebo_to_axi_7 = student_1,
              student_placebo_to_axi_15 = student_2,
              student_axi_7_to_axi_15 = student_3)
  
  final_df <- rbind(final_df, tmp)
  }
  return(final_df)
}
  


var_list <- c("peak_intensity", "area", "area_wash_in", "area_wash_out", "mean_transit_time", "mean_corrected", "time_to_peak")

statistical_results <- do.call(rbind, lapply(var_list, boxplots_tests, precoce))
statistical_results

```

```{r}

subset_for_plots <- merge(precoce, datas_pvs_long, by.x = c("mouse_id", "day"), by.y = c("souris", "day"))
subset_for_plots[, VEGFR_ratio := surface_VEGFR2/surface_tumeur]
subset_for_plots[, `:=`(date_greffe = as.Date(date_greffe, format = "%d/%m/%Y"),
                        date_inclusion = as.Date(date_inclusion, format = "%d/%m/%Y"),
                        date_sacrifice = as.Date(date_sacrifice, format = "%d/%m/%Y")
                        )]
subset_j11 <- subset_for_plots[day == "J11", ]


regression <- lm(mean_corrected ~VEGFR_ratio, subset_j11)
summary(regression)
regression$coefficients

g <- ggplot(subset_j11, aes(y = mean_corrected, x = VEGFR_ratio)) +
  geom_point(aes(color = Groupe))+
  geom_abline(intercept = regression$coefficients[1], slope = regression$coefficients[2])+
  scale_color_manual(values=c("#4373c5", "#ffc002", "#ed7d31"))+
  theme_minimal()
  
g
ggsave(file="../output_plots/table_correlation.svg", plot=g, width=10, height=8)

# cor.test(subset_j11$mean_corrected, subset_j11$VEGFR_ratio, method = "pearson")



g <- ggscatter(subset_for_plots[day == "J11", ], x = "VEGFR_ratio", y = "mean_corrected", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "VEGFR2 ratio", ylab = "Signal intensity (a.u)"
          )
g
ggsave(file="../output_plots/table_correlation_j11_pearson.svg", plot=g, width=10, height=8)
```


```{r}

pairs.panels(subset_for_plots[, .(peak_intensity,
                                  area, time_to_peak,
                                  area_wash_in, area_wash_out,
                                  mean_corrected, mean_base,
                                  mean_init, surface_histo,
                                  surface_echo, surface_tumeur,
                                  surface_VEGFR2,
                                  poids,
                                  volume,
                                  VEGFR_ratio)
                              ],
             smooth = TRUE,      # If TRUE, draws loess smooths
             scale = FALSE,      # If TRUE, scales the correlation text font
             density = TRUE,     # If TRUE, adds density plots and histograms
             ellipses = TRUE,    # If TRUE, draws ellipses
             method = "pearson", # Correlation method (also "spearman" or "kendall")
             pch = 21,           # pch symbol
             lm = FALSE,         # If TRUE, plots linear fit rather than the LOESS (smoothed) fit
             cor = TRUE,         # If TRUE, reports correlations
             jiggle = FALSE,     # If TRUE, data points are jittered
             factor = 2,         # Jittering factor
             hist.col = 4,       # Histograms color
             stars = TRUE,       # If TRUE, adds significance level with stars
             ci = TRUE)






```

